---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Weatherfunctions

<!-- badges: start -->
<!-- badges: end -->

The goal of Weatherfunctions is to support download of daily weather data from the German weather data service (DWD), to generate interpolations of weather data and to calculate some derived variables like potential evapotranspiration.

## Installation

You can install the development version of Weatherfunctions from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("AgronomyKiel/Weatherfunctions")
```

## Example

This is a basic example which shows you how to download a local copy of the historical and recent daily weather data from DWD and to interpolate them for a certain location defined by latitude, longitude and height above sea level.

```{r example,, message=FALSE, warning=FALSE}
rm(list = ls())
library(Weatherfunctions)
library(here)
library(fst)
library(dplyr)
library(tidyr)
library(sf)

# download meta data for "synoptic" weather stations from DWD
DWD_content <- getDWDStationList(historical = DWD_ftp_historical, recent = DWD_ftp_recent)

# download meta data for additional stations with precipitation data from DWD
DWDRain_content <- getDWDRainStationList(recent = DWDRain_ftp_recent, historical = DWDRain_ftp_historical)

# the package defines a standard directory for storing user specific data permanently 
outputDir <- Local_R_DWD

# save the meta data to local files
if (!dir.exists(outputDir)) {dir.create(outputDir, recursive = T)}
fn <- paste0(outputDir, fn_DWD_content)
save(DWD_content, file = fn)
fn <- paste0(outputDir, fn_DWDRain_content)
save(DWDRain_content, file = fn)


# The package supports two ways of saving locally the daily DWD weather data either as zip files or as fst files
# In general the DWD data are available as historic data (older than approx. 1 year) and as recent data (up to approx. 1 year), the latter are updated daily
# It is therefore convenient to save the historic data as zip and/or fst files locally
# This consumes some disk space and takes time for the first download, but it is much faster to access the data later on

# download the historical data to local zip files
# only download the historical data for the first time or once per year
#if (!dir.exists(LocalCopy_DWD_ftp_historical)) {dir.create(LocalCopy_DWD_ftp_historical, recursive = T)}
#Copy_DWD_ZipFiles(DWD_ftp_historical, LocalCopy_DWD_ftp_historical)


# download the recent data to local zip files is an option, not a must
#if (!dir.exists(LocalCopy_DWD_ftp_recent)) {dir.create(LocalCopy_DWD_ftp_recent, recursive = T)}
#Copy_DWD_ZipFiles(DWD_ftp_recent, LocalCopy_DWD_ftp_recent)


# update the data to local fst file from the local zip files (isloadnew = F) or directly from the DWD ftp server (isloadnew = T)
# weather data interpolation uses the fst format
#UpdateDWDData_to_fst(dataperiod = "historical", startdate = "1990-01-01", isloadnew = T)


# if the data have been stored locally, they can be read from the local fst files
fn_histDWD_data <- paste0(fn_histDWD_data_str, as.character(1990),".fst")
fn <- here(Local_R_DWD, fn_histDWD_data)
weather_historic <- read.fst(path = fn)

# select the columns that are in the core, i.e. essential data set and rename them
weather_historic_core <- weather_historic[,c("Stations_id", "Date",longNames_DWD_core)]
rm(weather_historic)

ThisYear <- as.numeric(format(Sys.Date(), "%Y"))
# take the data frame with meta data for recent stations to a separate variable
RecentRainStationList <- DWDRain_content$recent$stationlist

# update the recent data to local fst file from the local zip files (isloadnew = F) or directly from the DWD ftp server (isloadnew = T)
# if you want to update the recent data, set UpdateRecentDWDdata <- T
UpdateRecentDWDdata <- F
if (UpdateRecentDWDdata) {
  StartTime <- Sys.time()
  UpdateDWDData_to_fst(dataperiod = "recent",
                       isloadnew = T,
                       DWD_content = DWD_content,
                       startdate = paste0(as.character(ThisYear-1),"-01-01"))

  EndTime <- Sys.time()
  print(paste("Time for UpdateDWDData_to_fst:", EndTime - StartTime))
}

# the name of the file is generated
fn_recentDWD_data <- paste0("recent_weather_dat_", as.character(1990),".fst")
weather_recent <- read.fst(path = here(Local_R_DWD, fn_recentDWD_data))
weather_recent  <- weather_recent[,c("Stations_id","Date",longNames_DWD_core)]

# For estimation of weather data the coordinates and height asl of the location are needed
geoBreite <- 52.7306
geoLaenge <- 10.6298
Hoehe_m <- 85

# make location data frame
Location <- data.frame(Latitude=as.numeric(geoBreite), Longitude=as.numeric(geoLaenge))

# make an sf object from Location data frame
Location <- st_as_sf(Location, coords = c("Longitude", "Latitude")) %>%
  st_set_crs(value = "+proj=longlat +datum=WGS84")


### interpolate the historic weather data for the location

# retrieve the recent station list from the DWD content
# we need a list with the available stations for the historic data
stationlist <- DWD_content$historical$stationlist

# make an sf object from stationlist data frame
stationlist <- st_as_sf(stationlist, coords = c("geoLaenge", "geoBreite")) %>%
  st_set_crs(value = "+proj=longlat +datum=WGS84")

# add the distance to the location to the stationlist data frame
stationlist$Distance_m <- pmax(1,as.numeric(st_distance(stationlist, Location))) # minimum distance is 1 m because
stationlist$Distance_km <- as.numeric(format(stationlist$Distance_m /1000, digits = 3))
# remove the geometry column
stationlist$geometry <- NULL
stationlist <- as.data.frame(stationlist)
startdate <- "1990-01-01"

# The list of the additional stations with rain data is needed
RainStationList <- DWDRain_content$historical$stationlist

# select the 3 nearest stations with rain data, for historical data it is necessary to select more than one station
# because stations with rain data changed over time
RainStations_selected <- SelectStations (lat=geoBreite,
                                         long=geoLaenge,
                                         height_loc=Hoehe_m,
                                         stationlist = RainStationList,
                                         #DWDRain_content$recent$stationlist,
                                         minstations=3,
                                         max_stations = 3,
                                         radius=80000,
                                         startdate=startdate,
                                         max.Height.Distance_m=200)


# get the historic rain data for the selected stations
df_Rain <- GetRainData_selection(RainStations_selected,
                                 DWDRain_content,#$historical,
                                 repository=DWDRain_ftp_historical,
                                 startdate="1990-01-01")


# add the distance to the rain station to the data frame

df_Rain <- left_join(df_Rain, RainStations_selected[,c("Stations_id", "Distance_km")], by="Stations_id")

# interpolate the weather data for the location
IntPolDWDHistorical <- InterpolateFromDWD(df_DWD_core = weather_historic_core, stationlist = stationlist,
                                          geoBreite = geoBreite, geoLaenge = geoLaenge,
                                          max.Height.Distance_m=100, Hoehe_m = Hoehe_m, df_Rain = df_Rain,
                                          startdate = startdate)


IntPolDWDdataHistorical <- IntPolDWDHistorical$DWDdata

# remove the data for the current and previous year as they are in the recent data
ThisYear <- as.numeric(format(Sys.Date(), "%Y"))
FirstDayOfYear <- as.Date(paste0(ThisYear-1,"-01-01"))
ExcelFirstDayofYear <- as.numeric(FirstDayOfYear - as.Date("1899-12-30"))
IntPolDWDdataHistorical <- IntPolDWDdataHistorical %>% filter(Time < ExcelFirstDayofYear)

# Add longitude and latitude to the data frame
IntPolDWDdataHistorical$Longitude <- geoLaenge
IntPolDWDdataHistorical$Latitude <- geoBreite
```

The interpolation should deliver daily estimates of the weather data for the location:

```{r ExampleOutput}
summary(IntPolDWDdataHistorical)
```

The average distance for the estimation of the different weather variables can be calculated: 
```{r}
usedstations <- IntPolDWDHistorical$df.wf
usedstations$Distance <- 1/usedstations$InvDist
Diststations <- usedstations %>% group_by(variable) %>% dplyr::summarise(Distance = mean(Distance, na.rm = T))
```

The average distance for the estimation of the different weather variables is:
```{r}
Diststations
```

Now the same procedure for the recent data:

```{r}
# recent data #######

## retrieve the recent station list from the DWD content

DWDrecent <- DWD_content$recent
stationlist <- DWDrecent$stationlist

# make an sf object from stationlist data frame
stationlist <- st_as_sf(stationlist, coords = c("geoLaenge", "geoBreite")) %>%
  st_set_crs(value = "+proj=longlat +datum=WGS84")

# add the distance to the location to the stationlist data frame
stationlist$Distance_m <- pmax(1,as.numeric(st_distance(stationlist, Location))) # minimum distance is 1 m because
stationlist$Distance_km <- format(stationlist$Distance_m /1000, digits = 3)
# remove the geometry column
stationlist$geometry <- NULL
stationlist <- as.data.frame(stationlist)
startdate <- paste0(as.character(ThisYear-1),"-01-01")

# select the neart stations with rain data
RainStation_selected <- SelectStations (lat=geoBreite,
                                        long=geoLaenge,
                                        height_loc=Hoehe_m,
                                        stationlist = RecentRainStationList,
                                        #DWDRain_content$recent$stationlist,
                                        minstations=3,
                                        max_stations = 3,
                                        radius=80000,
                                        startdate=startdate,
                                        max.Height.Distance_m=200)


# get the rain data for the selected stations

df_Rain <- GetRainData_selection(RainStations_selected,
                                 DWDRain_content,
                                 repository=DWDRain_ftp_recent,
                                 startdate=startdate)


# add the distance to the rain station to the data frame

df_Rain <- left_join(df_Rain, RainStations_selected[,c("Stations_id", "Distance_km")], by="Stations_id")


# interpolate the recent weather data for the location
IntpolDWDrecent <- InterpolateFromDWD(df_DWD_core = weather_recent, stationlist = stationlist, geoBreite = geoBreite, geoLaenge = geoLaenge,
                                      max.Height.Distance_m=100, Hoehe_m = Hoehe_m, df_Rain = df_Rain,
                                      startdate = startdate)

IntPolDWDdatarecent <- IntpolDWDrecent$DWDdata



IntPolDWDdatarecent$Longitude <- geoLaenge
IntPolDWDdatarecent$Latitude <- geoBreite
```

Here the summary of the result of the interpolation procedure for the recent data:
```{r}
summary(IntPolDWDdatarecent)
```


```{r}
DWDdata <- rbind(IntPolDWDdataHistorical, IntPolDWDdatarecent)
```

Here the combinded data set:
```{r}
summary(DWDdata)
```


