---
title: "Weatherfunctions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Weatherfunctions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include=TRUE, warning=FALSE, message=FALSE}
rm(list=ls())
library(Weatherfunctions)
library(stringi)
library(RCurl)
library(here)
library(data.table)
library(dplyr)
library(fst)
library(sf)
library(dplyr)
library(purrr)
library(tidyr)
library(lubridate)
library(httr)
```


# Intro

The German Weather Service (DWD) supplies weather data in different spatial and time resolution and from different types of weather stations.
This package aims to a) assist in retrieving this data in a first step to a local copy of the data and b) in a second step to prepare the data for simulations and other purposes.
Thereby especially the estimation of global radiation data from sunshine hours is supported and interpolation for certain locations based on distance weighted use of the nearest weather station is possible.

## Structure of DWD ftp service

### Daily observations of weather data

The daily observations of weather data from the stations are divided into two sections, historical data which passed a quality control and recent data, not yet full controlled and usually containing the data of the actual and the data of the previous year.
Their actuality is usually one day back, so that the observations from the last day are available on the next day.
For convenience the link to these repositories and their local copies as well as already transformed data is stored in variable defined in the package.
For the historical data the variable "DWD_ftp_historical" is defined containing the link to `r DWD_ftp_historical` and for the recent data the variable "DWD_ftp_recent" with the link to `r DWD_ftp_recent`.
Also of interest are the directories with additional precipitation data and solar radiation data which are available only for a limited number of stations in Germany.
There is also a directory with selected long term observations from selected stations (timeseries_overview):

```         
├───kl
│   ├───historical
│   └───recent
├───more_precip
│   ├───historical
│   └───recent
├───solar
└───timeseries_overview
```

In order to reduce the number of file transfers when using these data it is convenient to make local copies of the DWD data.
The package defines a directory after installation were the downloads data are saved. It further  defines strings to the copies of recent and historical DWD data `r LocalCopy_DWD_ftp_historical`.  Depending on the local machine, the directory differs, for this machine the directory is:
`r  tools::R_user_dir("Weatherfunctions", which="data")` 
and can be retrieved by the command: tools::R_user_dir("Weatherfunctions", which="data"). 

## Structure of DWD data files

The data for the recent and historic observations are stored in ZIP files, one for each station, which contain a number of files with meta data and a data file (Example: produkt_klima_tag_20211123_20230526_00164.txt) were the filename consists of a string constant "produkt_klima_tag" two strings with the date of the first and the last observation day in the file in the format yyyymmdd and the Stations_id of the station as a 5 char string with leading zeros.

## Functions

### Get the station meta data (getDWDContent)

The function "getDWDContent" reads the content of the DWD ftp server and returns a data frame with the station meta data.
The function is used to get the station meta data for the historical and recent observations.

```{r getDWDContent, eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}

DWDhistorical <- getDWDContent(DWD_ftp_historical)
DWDrecent     <- getDWDContent(DWD_ftp_recent)
str(DWDhistorical)
```

The function returns a list with the following elements:

- "stationlist" a data frame with the station meta data
- "filelist" a string array with the filenames of the meta data files 
- "ziplist" a list of the ZIP files for the stations
- "zipID" a list of the station IDs

### Meta data for additional precipitation stations

For the additional data from stations measuring precipitation data the function "getDWDContent" has to be called with the respective parameters.

```{r getDWDContent_precipitation, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
DWDRain_content <- getDWDRainStationList(recent = DWDRain_ftp_recent, historical = DWDRain_ftp_historical)

outputDir <- Local_R_DWD 
if (!dir.exists(outputDir)) {dir.create(outputDir, recursive = T)}
fn <- paste0(outputDir, fn_DWDRain_content)
save(DWDRain_content, file = fn)
nHistoricalStationlist <- nrow(DWDRain_content$historical$stationlist)
nRecentStationlist <- nrow(DWDRain_content$recent$stationlist)

```

The DWD records for historical observations consists of data r nHistoricalStationlist  different historical stations (which changed over time) and for the recent observations of values of r nRecentStationlist  stations.


### Retrieving the combined (recent & historic) information (getDWDStationList)

The function getDWDStationList retrieves the information (meta data) for both, recent and historic directories.
The function returns a compiled data frame containing all station parameters ("stations") as well as the lists of the historical and recent weather station information the function "getDWDContent" returns (see section above).

```{r getDWDStationList, eval=FALSE, message=FALSE, warning=FALSE}
DWD_content <- getDWDStationList(historical = DWD_ftp_historical, recent = DWD_ftp_recent)
str(DWD_content)
```

### Make a local copy of DWD observation data (Copy_DWD_ZipFiles)

For the file transfer the package holds the function "Copy_DWD_ZipFiles" for downloading the files to a local directory.
For the recent data the package defines the path `r LocalCopy_DWD_ftp_recent` and for the historical data the path `r LocalCopy_DWD_ftp_historical`.

There are 500+ recent stations available and 1000+ data for recent & historic stations.
So the download takes some time.

For a total download of the recent and historical data the function "Copy_DWD_ZipFiles" the function has to be called with the respective parameters twice.

```{r Copy_DWD_ZipFiles, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}

# download the recent data
if (!dir.exists(LocalCopy_DWD_ftp_recent)) {dir.create(LocalCopy_DWD_ftp_recent, recursive = T)}
Copy_DWD_ZipFiles(DWD_ftp_recent, LocalCopy_DWD_ftp_recent)

# download the historical data
if (!dir.exists(LocalCopy_DWD_ftp_historical)) {dir.create(LocalCopy_DWD_ftp_historical, recursive = T)}
Copy_DWD_ZipFiles(DWD_ftp_historical, LocalCopy_DWD_ftp_historical)



```
For further processing of the data and to save disk space the data can be saved in a more compact format using the function "UpdateDWDData_to_fst". The parameter isloadnew is used to load the data from the DWD ftp server, if FALSE the local zip-files will be used, if available. Again this process is time consuming, as the data is read from the zip files and saved in a fst file. The parameter startdate is used to define the start date of the data to be processed. The file size may still be large, the data in fst format for the historical data from 1970 on is about 1 GB.



```{r eval=FALSE, include=FALSE}
UpdateDWDData_to_fst(dataperiod = "historical", startdate = "1990-01-01", isloadnew = F)
UpdateDWDData_to_fst(dataperiod = "recent", startdate = "1990-01-01", isloadnew = F)

```



## Download DWD data

The package Weatherfunctions provides functions to download the data from the DWD ftp server or its local copy.

### Data for a single weather station recent or historical (getsingleDWDWeather)

The data from the DWD repository or its local copy can be downloaded either from the historic or recent data repositories which can be the DWD ftp service or its local copy using the function getsingleDWDWeather.

```{r getsingleDWDWeather, eval=FALSE, message=FALSE, warning=FALSE}
selstation <-  as.character(DWD_content$recent$zipID[3])
recent_data <- NULL
recent_data <- getsingleDWDWeather(station = selstation, ziplist = DWD_content$recent$ziplist,
                            repository = DWD_ftp_recent, quiet = F)
```


```{r, eval=FALSE}
selstation <-  as.character(DWDRain_content$recent$zipID[3])
test <- getsingleDWDRain(station = selstation, ziplist = DWDRain_content$recent$ziplist,
                            repository = DWDRain_ftp_recent, quiet = F)
```

```{r getsingleDWDWeatheroutput, message=FALSE, warning=FALSE, eval=FALSE}
test <- RenameDWDRain(test)
head(test)

```
In addition to the data from produkt_klima_tag_YYYYMMDD_YYYYMMDD_nnnnn.txt, the measurement height of wind speed was retrieved from the metadata file of the selected weather stations and a Date column is added in R format.

Note that the windspeed data is given for a height of 10 m above ground level. If wind speed was measured at a different height the correction is made with the function "windheight".

$$u_o = u_i \cdot \frac{ log(67.8 \cdot z_o - 5.42)} { log(67.8 * z_i - 5.42)}$$

The downloaded data have column names according to the code of DWD, for which in a first step also readable names are defined.
in the data frame df_names_DWD.

```{r DWDdata, eval=FALSE,}
df_names_DWD

```



### Download combined data (getcombinedDWDWeather)

For downloading the combination of historic and recent weather data of a particular station the function "getcombinedDWDWeather" can be used.
It also adds station information to the returned data frame.

```{r getcombinedDWDWeather, eval=FALSE}
selstation <-  as.character(DWD_content$recent$zipID[3])
combined_data <- NULL
combined_data <- getcombinedDWDWeather(DWD_content = DWD_content, station = selstation,
                              recent_repository = DWD_ftp_recent, historical_repository = DWD_ftp_historical, local = F)
combined_data <- RenameDWDWeather(combined_data)
str(combined_data)

```

## Add radiation data (getRadWeather)

The function getRadWeather uses the DWD-data in long name format and adds several derived parameters.

```{r getRadWeather, eval=FALSE}
RadWeather <- NULL
RadWeather <- getRadWeather(combined_data)
str(RadWeather)
```

The additional columns are DOY: day of year, Dayl: daylength in hours, Angot: maximum possible radiation outside atmosphere [MJ.m-2.d-1], RelSun: the relative sunshine hours on that day, Monat: the calendar month of the day, , EstGlobRad the estimated global radiation of the day, ExcelTime: the day of year in Excel format, VP: the average vapour pressure on the day [mBar], the saturation deficit of the air and Ra_Int, the radiation intensity in [W.m-2].
The functions uses other functions from the package and external functions to calculate some of these parameters.
*Calcsolar (dayofyear, latitude)* calculates the daylength and the angot value.
The function *relrad(RelSun, Month)* calculates the relative radiation compared to the angot values of that day of year.
The function is based on an empirical regression between the relative radiation the relative sunshine duration for different months in Germany based on DWD data.



## Saving data to local files in fst-format (UpdateDWDData_to_fst)

In order to save the downloaded data in a more compact and pre-computed format the function UpdateDWDData_to_fst can be used.The parameter '*isloadnew*' is used to load the data from the DWD ftp server, if FALSE the local zip-files will be used, if available.
This again is a time consuming process, as the data is read from the zip files and saved in a fst file.
For further analysis it is often sufficient to download the historical data only to the fst file, as these data are updated only once a year.



```{r UpdateDWDData_to_fst, message=FALSE, warning=FALSE, eval=FALSE}

UpdateDWDData_to_fst(dataperiod = "recent", startdate = "1990-01-01", isloadnew = T)

```

A saved file can be read using the function read.fst from the package fst.
 
```{r, eval=FALSE}
fn <- "recent_weather_dat_1990.fst"
path <- paste(Local_R_DWD, fn, sep="/")
if (file.exists(path)){
  weather_recent <- read.fst(path = path)}
```

The function UpdateDWDData_to_fst can also be used to update the historical data. This is done by setting the parameter dataperiod to "historical" and the startdate to the first date of the data to be downloaded. This process is time consuming and should be done only once a year.

```{r UpdateDWDData_to_fst_historical, message=FALSE, warning=FALSE, eval=FALSE}
UpdateDWDData_to_fst(dataperiod = "historical", startdate = "1990-01-01", isloadnew = T)

```


## Interpolation of weather data for a selected location

The algorithm is basically based on the approach to select for every day and for every weather parameter the values of the 3 nearest observations and to interpolate between those values according to inverse distance weigthing.

Additionally the rainfall data of the nearest weather station including the additional rainfall stations are used.

The function InterpolateFromDWD is used to interpolate the weather data for a selected location. The procedure is split in two parts, first the interpolation of the historical data and second the interpolation of the recent data, finally the data is combined in one data frame.

The retrieval of the historical and - to a lesser extent - the recent data is time consuming, as the data for each station has to be downloaded as zip files  and extracted. It is therefore more efficient to pre download the data. The package supports two different approaches, the first is to  a fst file for further use. Here the the fst approach is used.

### Retrieval of the weather data for interpolation

```{r LoadDataForInterpolation, message=FALSE, warning=FALSE, eval=FALSE}


# load historic weather data from local fst file ####
# the filename is constructed from two strings for the actual path of the project
# and a filename which are both pre defined.

fn_histDWD_data <- paste0(fn_histDWD_data_str, as.character(1990),".fst")

fn <- here(Local_R_DWD, fn_histDWD_data)
weather_historic <- read.fst(path = fn)

# select the columns that are in the core, i.e. essential data set and rename them
weather_historic_core <- weather_historic[,c("Stations_id", "Date",longNames_DWD_core)]
rm(weather_historic)

# update the local weather data for the recent period ######################
# and store them in a fst file
# this has to be done only once per day
# the data is stored in a fst file
# and takes up to 10-15 minutes

ThisYear <- as.numeric(format(Sys.Date(), "%Y"))
# take the data frame with meta data for recent stations to a separate variable
RecentRainStationList <- DWDRain_content$recent$stationlist

UpdateRecentDWDdata <- F
if (UpdateRecentDWDdata) {
  StartTime <- Sys.time()
  UpdateDWDData_to_fst(dataperiod = "recent", 
                       isloadnew = T,
                       DWD_content = DWD_content,
                       startdate = paste0(as.character(ThisYear-1),"-01-01"))
  
  EndTime <- Sys.time()
  print(paste("Time for UpdateDWDData_to_fst:", EndTime - StartTime))
}


fn_recentDWD_data <- paste0("recent_weather_dat_", as.character(1990),".fst")
weather_recent <- read.fst(path = here(Local_R_DWD, fn_recentDWD_data))
weather_recent  <- weather_recent[,c("Stations_id","Date",longNames_DWD_core)] 

# set the location of the station  
  geoBreite <- 52.7306
  geoLaenge <- 10.6298
  Hoehe_m <- 85
  
  # make location data frame  
  Location <- data.frame(Latitude=as.numeric(geoBreite), Longitude=as.numeric(geoLaenge))
  
  # make an sf object from Location data frame
  Location <- st_as_sf(Location, coords = c("Longitude", "Latitude")) %>%
    st_set_crs(value = "+proj=longlat +datum=WGS84")

```

The historic weather data contains the data for the years 1990 to the previous year and consists of r nrow(weather_historic_core) records.
The recent weather data contains the data for the years r ThisYear-1 to the current year and consists of r nrow(weather_recent) records.


### Interpolation of the historic weather data for the location

In a next step the historic weather data is interpolated for the location.
If all gets ok then the data frame should have no NA values.

```{r InterpolateHistoricData, message=FALSE, warning=FALSE, eval=FALSE}

  
### interpolate the historic weather data for the location  
  
  # retrieve the recent station list from the DWD content
  stationlist <- DWDhistorical$stationlist

  # make an sf object from stationlist data frame
  stationlist <- st_as_sf(stationlist, coords = c("geoLaenge", "geoBreite")) %>%
    st_set_crs(value = "+proj=longlat +datum=WGS84")
  
  # add the distance to the location to the stationlist data frame
  stationlist$Distance_m <- pmax(1,as.numeric(st_distance(stationlist, Location))) # minimum distance is 1 m because
  stationlist$Distance_km <- as.numeric(format(stationlist$Distance_m /1000, digits = 3))
  # remove the geometry column
  stationlist$geometry <- NULL
  stationlist <- as.data.frame(stationlist)
  startdate <- "1990-01-01"
  RainStationList <- DWDRain_content$historical$stationlist
  # select the nearest stations with rain data, for historical data it is necessary to select more than one station
  # because stations with rain data changed over time
  RainStations_selected <- SelectStations (lat=geoBreite,
                                          long=geoLaenge,
                                          height_loc=Hoehe_m, 
                                          stationlist = RainStationList,
                                          #DWDRain_content$recent$stationlist,
                                          minstations=3,
                                          max_stations = 3,
                                          radius=80000,
                                          startdate=startdate,
                                          max.Height.Distance_m=200)
  
  
  # get the rain data for the selected stations
  
  df_Rain <- GetRainData_selection(RainStations_selected,
                                     DWDRain_content,#$historical,
                                     repository=DWDRain_ftp_historical,
                                     startdate="1990-01-01") 
  
  
  # add the distance to the rain station to the data frame

  df_Rain <- left_join(df_Rain, RainStations_selected[,c("Stations_id", "Distance_km")], by="Stations_id")
  
  # interpolate the weather data for the location
  IntPolDWDHistorical <- InterpolateFromDWD(df_DWD_core = weather_historic_core, stationlist = stationlist, 
                                          geoBreite = geoBreite, geoLaenge = geoLaenge, 
                                max.Height.Distance_m=100, Hoehe_m = Hoehe_m, df_Rain = df_Rain, 
                                startdate = startdate) 
  IntPolDWDdataHistorical <- IntPolDWDHistorical$DWDdata
  
  # remove the data for the current and previous year as they are in the recent data
  ThisYear <- as.numeric(format(Sys.Date(), "%Y"))
  FirstDayOfYear <- as.Date(paste0(ThisYear-1,"-01-01"))
  ExcelFirstDayofYear <- as.numeric(FirstDayOfYear - as.Date("1899-12-30"))
  IntPolDWDdataHistorical <- IntPolDWDdataHistorical %>% filter(Time < ExcelFirstDayofYear)

  # Add longitude and latitude to the data frame
  IntPolDWDdataHistorical$Longitude <- geoLaenge
  IntPolDWDdataHistorical$Latitude <- geoBreite
  summary(IntPolDWDdataHistorical)
  
  usedstations <- IntPolDWDHistorical$df.wf
  usedstations$Distance <- 1/usedstations$InvDist
  Diststations <- usedstations %>% group_by(variable) %>% dplyr::summarise(Distance = mean(Distance, na.rm = T))
  Diststations


```


### Interpolation of the recent weather data for the location

In a next step the recent weather data is interpolated for the location.
Here, the data for the most actual date(s) may be incomplete.

```{r InterpolateRecentData, message=FALSE, warning=FALSE, eval=FALSE}

  
# recent data #######

  ## retrieve the recent station list from the DWD content
  
  stationlist <- DWDrecent$stationlist

  # make an sf object from stationlist data frame
  stationlist <- st_as_sf(stationlist, coords = c("geoLaenge", "geoBreite")) %>%
    st_set_crs(value = "+proj=longlat +datum=WGS84")
  
  # add the distance to the location to the stationlist data frame
  stationlist$Distance_m <- pmax(1,as.numeric(st_distance(stationlist, Location))) # minimum distance is 1 m because
  stationlist$Distance_km <- format(stationlist$Distance_m /1000, digits = 3)
  # remove the geometry column
  stationlist$geometry <- NULL
  stationlist <- as.data.frame(stationlist)
  startdate <- paste0(as.character(ThisYear-1),"-01-01")
  
  # select the neart stations with rain data
  RainStation_selected <- SelectStations (lat=geoBreite,
                                          long=geoLaenge,
                                          height_loc=Hoehe_m, 
                                          stationlist = RecentRainStationList,
                                          #DWDRain_content$recent$stationlist,
                                          minstations=3,
                                          max_stations = 3,
                                          radius=80000,
                                          startdate=startdate,
                                          max.Height.Distance_m=200)
  
  
  # get the rain data for the selected stations
  
  df_Rain <- GetRainData_selection(RainStation_selected,
                                     DWDRain_content,
                                     repository=DWDRain_ftp_recent,
                                     startdate="2023-01-01") 
  
  
  # add the distance to the rain station to the data frame

  df_Rain <- left_join(df_Rain, RainStations_selected[,c("Stations_id", "Distance_km")], by="Stations_id")


  #
  IntpolDWDrecent <- InterpolateFromDWD(df_DWD_core = weather_recent, stationlist = stationlist, geoBreite = geoBreite, geoLaenge = geoLaenge, 
                                max.Height.Distance_m=100, Hoehe_m = Hoehe_m, df_Rain = df_Rain, 
                                startdate = startdate) 
  
  IntPolDWDdatarecent <- IntpolDWDrecent$DWDdata
 
  
  
  IntPolDWDdatarecent$Longitude <- geoLaenge
  IntPolDWDdatarecent$Latitude <- geoBreite
  
  summary(IntPolDWDdatarecent)

```

## Combine the historical and recent data

The data from the historical and recent data is combined in one data frame.

```{r, eval=FALSE}
  
  DWDdata <- rbind(IntPolDWDdataHistorical, IntPolDWDdatarecent)
  summary(DWDdata)

```

The combined data frame contains r nrow(DWDdata) records and starts at r as.Date(min(DWDdata$Time),origin = "1899-12-30" )` and ends at 
r as.Date(max(DWDdata$Time),origin = "1899-12-30" ).




